---
title: "mosaicGG Function"
author: "Sean F. Larsen"
date: "May 22, 2019"
output: html_document
---

## mosaicGG Function

A function is created to allow the plotting of other datasets into the same charts.
This Code is a set-up of that function.  First the code will read the csv file, set-up the dataframe, correct the column names, and then rewrite the data into a new csv file adult_update.csv

```{r echo=TRUE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(ggthemes)

adult <- read.csv(file="adult.csv", header=TRUE, sep=",")
adult <- data.frame(adult$RBMI, adult$BMI_P, adult$RACEHPR2, adult$SRSEX, adult$SRAGE_P, adult$MARIT2, adult$AB1, adult$ASTCUR, adult$AB51, adult$POVLL)
colnames(adult) = c("RBMI", "BMI_P", "RACEHPR2", "SRSEX", "SRAGE_P", "MARIT2", "AB1", "ASTCUR", "AB51", "POVLL")
head(adult)
```

## Set-Up the Function

This code recreates the mosaicGG function from the exercise.

```{r echo= TRUE}
mosaicGG <- function(data, X, FILL) {
  # Proportions in raw data
  DF <- as.data.frame.matrix(table(data[[X]], data[[FILL]]))
  DF$groupSum <- rowSums(DF)
  DF$xmax <- cumsum(DF$groupSum)
  DF$xmin <- DF$xmax - DF$groupSum
  DF$X <- row.names(DF)
  DF$groupSum <- NULL
  DF_melted <- melt(DF, id = c("X", "xmin", "xmax"), variable.name = "FILL")
  DF_melted <- DF_melted %>%
    group_by(X) %>%
    mutate(ymax = cumsum(value/sum(value)),
           ymin = ymax - value/sum(value))

  # Chi-sq test
  results <- chisq.test(table(data[[FILL]], data[[X]])) # fill and then x
  resid <- melt(results$residuals)
  names(resid) <- c("FILL", "X", "residual")

  # Merge data
  DF_all <- merge(DF_melted, resid)

  # Positions for labels
  DF_all$xposn <- DF_all$xmin + (DF_all$xmax - DF_all$xmin)/2
  index <- DF_all$xmax == max(DF_all$xmax)
  DF_all$yposn <- DF_all$ymin[index] + (DF_all$ymax[index] - DF_all$ymin[index])/2

  # Plot
  g <- ggplot(DF_all, aes(ymin = ymin,  ymax = ymax, xmin = xmin,
                          xmax = xmax, fill = residual)) +
  geom_rect(col = "white") +
  geom_text(aes(x = xposn, label = X),
            y = 1, size = 3, angle = 90, hjust = 1, show.legend = FALSE) +
  geom_text(aes(x = max(xmax),  y = yposn, label = FILL),
            size = 3, hjust = 1, show.legend = FALSE) +
  scale_fill_gradient2("Residuals") +
  scale_x_continuous("Individuals", expand = c(0,0)) +
  scale_y_continuous("Proportion", expand = c(0,0)) +
  theme_tufte() +
  theme(legend.position = "bottom")
  print(g)
}
```

## Generalization

Now the fuction is used to plot with various datasets, starting with the adult dataset already created.

```{r echo= TRUE}
# BMI described by age (as previously seen)
mosaicGG(adult, X = "SRAGE_P", FILL = "RBMI")

# Poverty described by age
mosaicGG(adult, X = "SRAGE_P", FILL = "POVLL")

# mtcars: am described by cyl
mosaicGG(mtcars, "cyl", "am")

# Vocab: vocabulary described by education
library(carData)
mosaicGG(Vocab, "education", "vocabulary")
```